---
title: "add-term1-term2-example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{add-term1-term2-example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(metalite.ae)
```


```{r}
meta <- meta_ae_example()
# Check if AEOSI exists
"AEOSI" %in% names(meta$data_observation)
```

```{r}
# Access the correct observation data
adae_modified <- meta$data_observation

# Create AEOSI based on specific adverse event terms of special interest
special_interest_terms <- c(
  "DIARRHOEA", "NAUSEA", "VOMITING", "HEADACHE", 
  "DIZZINESS", "SOMNOLENCE", "TREMOR"
)

adae_modified$AEOSI <- ifelse(
  adae_modified$AEDECOD %in% special_interest_terms, 
  "Y", 
  "N"
)

# Update the meta object with the correct path
meta$data_observation <- adae_modified

```



```{r}
# Check if AEOSI exists
"AEOSI" %in% names(meta$data_observation)

# View the distribution
table(meta$data_observation$AEOSI)

```


```{r}
param_info <- metalite::collect_adam_mapping(meta, "aeosi0rel")
print(param_info$term1)
print(param_info$term2)
```

```{r}
new_plan <- plan(
  analysis = "ae_summary", population = "apat",
  observation = c("wk12", "wk24"), parameter = "any;rel;ser"
) |>
  add_plan(
    analysis = "ae_specific", population = "apat",
    observation = c("wk12", "wk24"),
    parameter = c("any", "aeosi", "rel", "ser", "dtc0rel", "aeosi0rel")  # Added aeosi0rel
  ) |>
  add_plan(
    analysis = "ae_listing", population = "apat",
    observation = c("wk12", "wk24"), parameter = c("any", "rel", "ser")
  ) |>
  add_plan(
    analysis = "ae_exp_adj", population = "apat",
    observation = c("wk12", "wk24"), parameter = "any;rel;ser"
  )
# Update the meta object with new plan
meta <- meta |>
  define_plan(plan = new_plan) |>
  define_parameter(
    name = "aeosi0rel",
    subset = quote(AEOSI == "Y"),
    var = "AEDECOD",
    soc = "AEBODSYS",
    term1 = " ",
    term2 = " ",
    label = "adverse events of special interest"
  ) |>
  meta_build()
```


```{r}
param_info <- metalite::collect_adam_mapping(meta, "aeosi0rel")
print(param_info$term1)
print(param_info$term2)
```

```{r, message = FALSE}
outdata <- prepare_ae_specific(
  meta,
  population = "apat",
  observation = "wk12",
  parameter = "aeosi0rel"
)
```



```{r}
outdata
```

## RTF tables

The last step is to prepare the RTF table using `tlf_ae_summary()`.

```{r}
outdata |>
  format_ae_specific() |>
  tlf_ae_specific(
    meddra_version = "24.0",
    source = "Source:  [CDISCpilot: adam-adsl; adae]",
    analysis = "ae_specific", # Provide analysis type defined in meta$analysis,
    orientation = "landscape",
    col_rel_width = c(6, rep(1, 8)),
    path_outtable = "rtf/ae0specific0aeosi0missing.rtf"
  )
```


```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("pdf/ae0specific0aeosi0missing.pdf")
```



```{r}
outdata$group
```

The row is indexed according to the order of `outdata$name`.

```{r}
head(data.frame(outdata$order, outdata$name))
```

- `n_pop`: number of participants in population.

```{r}
outdata$n_pop
```

- `n`: number of subjects with AE.

```{r}
head(outdata$n)
```

- `prop`: proportion of subjects with AE.

```{r}
head(outdata$prop)
```

- `diff`: risk difference compared with the `reference_group`.

```{r}
head(outdata$diff)
```

## Format output

Once the raw analysis results are obtained,
the `format_ae_specific()` function can be employed to prepare the outdata,
ensuring its compatibility with production-ready RTF tables.

```{r}
tbl <- outdata |> format_ae_specific()
head(tbl$tbl)
```

### Additional statistics

By using the `display` argument,
we can choose specific statistics to include.
For instance, we have the option to incorporate the risk difference.

```{r}
tbl <- outdata |> format_ae_specific(display = c("n", "prop", "diff"))
head(tbl$tbl)
```

To perform advanced analysis, the `extend_ae_specific_inference()` function
is utilized.
For instance, we can incorporate a 95% confidence interval based on the
Miettinen and Nurminen (M&N) method.
Further information regarding the M&N method can be found in the
[rate compare vignette](https://merck.github.io/metalite.ae/articles/rate-compare.html).

```{r}
tbl <- outdata |>
  extend_ae_specific_inference() |>
  format_ae_specific(display = c("n", "prop", "diff", "diff_ci"))
head(tbl$tbl)
```

We can use `extend_ae_specific_duration()` to add average duration of AE.

```{r}
tbl <- outdata |>
  extend_ae_specific_duration(duration_var = "ADURN") |>
  format_ae_specific(display = c("n", "prop", "dur"))

head(tbl$tbl)
```

We can use `extend_ae_specific_events()` to add number of AE and/or average of it per subject.

```{r}
tbl <- outdata |>
  extend_ae_specific_events() |>
  format_ae_specific(display = c("n", "prop", "events_count", "events_avg"))

head(tbl$tbl)
```

We can use `filter_method` and `filter_criteria` parameters to filter information based on the specified criteria:

- `filter_method`: A character value to specify how to filter rows (by `count` or `percent`).
    - `count`: Filter based on participant count.
    - `percent`: Filter based on percent incidence.
- `filter_criteria`: A numeric value to display rows where at least one therapy group has:
    - a percent incidence or participant count greater than or equal to the specified value.
    - If `filter_method` is `percent`, the value should be between 0 and 100.
    - If `filter_method` is `count`, the value should be greater than 0.


```{r}
tbl <- outdata |>
  extend_ae_specific_events() |>
  format_ae_specific(
    display = c("n", "prop", "events_count", "events_avg"),
    filter_method = "percent",
    filter_criteria = 6
  )

head(tbl$tbl)
```
In results above, rows having any one of "prop_x" values are greater than 6 get kept in the output.


We can use `sort_order` and `sort_column` parameters to sort results based on the specified criteria:

- `sort_order` A character value to specify sorting order:
    - `alphabetical`: Sort by alphabetical order.
    - `count_des`: Sort by count in descending order.
    - `count_asc`: Sort by count in ascending order.
- `sort_column A` character value of `group` in `outdata` used to sort a table with.

```{r}
tbl <- outdata |>
  extend_ae_specific_events() |>
  format_ae_specific(
    display = c("n", "prop", "events_count", "events_avg"),
    sort_order = c("count_des"),
    sort_column = c("Placebo")
  )

head(tbl$tbl)
```

### Mock data preparation

The `mock` argument facilitates the creation of a mock table with ease.

Please note that the intention of the `mock` argument is not to provide
an all-encompassing mock table template.
Instead, it serves as a convenient method to assist users in generating
a mock table that closely resembles the desired output layout.
To develop a more versatile mock table generation tool, further efforts
are necessary.
This could potentially involve the creation of a dedicated mock table
generation package or similar solutions.

```{r}
tbl <- outdata |> format_ae_specific(mock = TRUE)
head(tbl$tbl)
```

## RTF tables

The last step is to prepare the RTF table using `tlf_ae_summary()`.

```{r}
outdata |>
  format_ae_specific() |>
  tlf_ae_specific(
    meddra_version = "24.0",
    source = "Source:  [CDISCpilot: adam-adsl; adae]",
    analysis = "ae_specific", # Provide analysis type defined in meta$analysis,
    orientation = "landscape",
    col_rel_width = c(6, rep(1, 8)),
    path_outtable = "rtf/ae0specific0aeosi0missing.rtf"
  )
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("pdf/ae0specific1.pdf")
```