---
title: "AE Listing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AE Listing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  out.width = "100%",
  dpi = 150
)
```

```{r}
library(metalite.ae)
library(metalite)
```

## Overview

The purpose of this tutorial is to create production ready AE listing.

The AE listing provides details about the interested adverse events information.
With a metadata object created by metalite,
there are two required functions to create AE listing tables using metalite.ae:

- `prepare_ae_listing()`: prepare listing datasets.
- `tlf_ae_listing()`: transfer output dataset to RTF table.

An example output:

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("pdf/ae0listing0ser0wk12.pdf")
```

## Example data

In metalite.ae, we created an example dataset using ADSL and ADAE datasets from the metalite package.

```{r}
meta <- meta_ae_example()
meta
```

The same metadata structure is used to support all analyses examples in metalite.ae.
More details can be found on the [metalite package website](https://merck.github.io/metalite/articles/metalite.html).

## Analysis preparation

The `prepare_ae_listing()` function is used to calculate statistics required for
AE listing using pre-specified keywords in `meta`.
The input of the function is a `meta` object created by metalite.
The output of the function is an `outdata` object that contains a list of analysis raw datasets.

```{r}
tbl <- prepare_ae_listing(
  meta,
  analysis = "ae_listing",
  population = "apat",
  observation = "wk12",
  parameter = "ser"
)
head(tbl$tbl)
head(tbl$col_name)
```

## RTF tables

The last step is to prepare the RTF table using `tlf_ae_listing()`.

```{r}
footnote <- c(
  "Related: Investigator-assessed relationship of the adverse event to study medication. Y = RELATED, N = NOT RELATED",
  "Action Taken: Discontinued = DRUG WITHDRAWN, Interrupted = DRUG INTERRUPTED, Reduced = DOSE REDUCED, Increased = DOSE INCREASED, None = DOSE NOT CHANGED, N/A = NOT APPLICABLE.",
  "Outcome: Resolved = RECOVERED/RESOLVED, Resolving = RECOVERING/RESOLVING, Sequelae = RECOVERED/RESOLVED WITH SEQUELAE, Not resolved = NOT RECOVERED/NOT RESOLVED.",
  "Adverse event terms are from MedDRA Version 25.0."
)
```

```{r}
tbl |> tlf_ae_listing(
  footnotes = footnote,
  orientation = "portrait",
  source = "Source:  [CDISCpilot: adam-adsl; adae]",
  path_outtable = "rtf/ae0listing0ser0wk12.rtf",
  path_outdata = NULL
)
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("pdf/ae0listing0ser0wk12.pdf")
```
