---
title: "AE Listing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AE Listing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  out.width = "100%",
  dpi = 150
)
```

```{r}
library(metalite.ae)
```

## Overview

The objective of this vignette is to generate a static adverse events (AE) listing table with serious AE.

The AE listing offers comprehensive information on the desired adverse events.
There are two essential functions for constructing AE listing tables with metalite.ae:

- `prepare_ae_listing()`: prepare AE listing datasets.
- `tlf_ae_listing()`: transfer output datasets to RTF files.


## Step 1: Define metadata

The example uses ADSL and ADAE data from the
[forestly](https://merck.github.io/forestly/) package.

```{r}
adsl <- forestly::forestly_adsl
adae <- forestly::forestly_adae

adsl$TRT01A <- factor(
  adsl$TRT01A,
  levels = c("Xanomeline Low Dose", "Placebo"),
  labels = c("Low Dose", "Placebo")
)
adae$TRTA <- factor(
  adae$TRTA,
  levels = c("Xanomeline Low Dose", "Placebo"),
  labels = c("Low Dose", "Placebo")
)

analysis_plan <- metalite::plan(
  analysis = "ae_listing",
  population = "apat",
  observation = "wk12",
  parameter = "ser"
)

meta <- metalite::meta_adam(observation = adae, population = adsl) |>
  metalite::define_plan(analysis_plan) |>
  metalite::define_population(
    name = "apat",
    var = c(
      "USUBJID", "SAFFL", "TRT01A", "TRTDUR",
      "SITEID", "SEX", "RACE", "AGE"
    ),
    group = "TRT01A",
    subset = SAFFL == "Y",
    label = "All Participants as Treated"
  ) |>
  metalite::define_observation(
    name = "wk12",
    var = c(
      "USUBJID", "SAFFL", "TRTA", "AEDECOD", "AEBODSYS", "AEREL",
      "AESER", "AEOUT", "AEACN", "AESDTH", "ASTDT", "AENDT"
    ),
    group = "TRTA",
    subset = SAFFL == "Y",
    label = "Weeks 0 to 12"
  ) |>
  metalite::define_parameter(
    name = "ser",
    term1 = "Serious",
    term2 = "",
    subset = AESER == "Y",
    var = "AEDECOD",
    soc = "AEBODSYS",
    label = "Serious AEs"
  ) |>
  metalite::define_analysis(
    name = "ae_listing",
    var_name = c(
      "USUBJID", "ASTDY", "AEDECOD", "ADURN",
      "AESEV", "AESER", "AEREL", "AEOUT"
    ),
    group_by = c("USUBJID", "ASTDY"),
    page_by = "TRTA"
  ) |>
  metalite::meta_build()
```

## Step 2: Generate the AE listing table

The function `prepare_ae_listing()` is used to create a dataset for
AE listing by utilizing predefined keywords specified
in the example data `meta`.

The last step is to prepare the RTF table using `tlf_ae_listing()`.

```{r}
footnote <- c(
  "Related: Investigator-assessed relationship of the adverse event to study medication. Y = RELATED, N = NOT RELATED",
  "Action Taken: Discontinued = DRUG WITHDRAWN, Interrupted = DRUG INTERRUPTED, Reduced = DOSE REDUCED, Increased = DOSE INCREASED, None = DOSE NOT CHANGED, N/A = NOT APPLICABLE.",
  "Outcome: Resolved = RECOVERED/RESOLVED, Resolving = RECOVERING/RESOLVING, Sequelae = RECOVERED/RESOLVED WITH SEQUELAE, Not resolved = NOT RECOVERED/NOT RESOLVED.",
  "Adverse event terms are from MedDRA Version 25.0."
)

prepare_ae_listing(
  meta,
  analysis = "ae_listing",
  population = "apat",
  observation = "wk12",
  parameter = "ser"
  ) |>
  tlf_ae_listing(
    footnotes = footnote,
    orientation = "portrait",
    source = "Source:  [CDISCpilot: adam-adsl; adae]",
    analysis = "ae_listing", # Provide analysis type defined in meta$analysis
    path_outtable = "./vignettes/rtf/ae0listing1.rtf")
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("pdf/ae0listing1.pdf")
```
