---
title: "Generate a Static AE Listing Table in RTF format"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generate a Static AE Listing Table in RTF format}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  out.width = "100%",
  dpi = 150
)
```

```{r}
library(metalite.ae)
```

## Overview

This vignette demonstrates how to generate a static adverse event (AE) listing
focused on serious AEs in **RTF** format.

The listing presents participant-level details for adverse events of interest.
Two functions support the workflow:

- `prepare_ae_listing()` prepares the listing dataset.
- `tlf_ae_listing()` creates the RTF table.

## Step 1: Define metadata

The example uses ADSL and ADAE data from the
[forestly](https://merck.github.io/forestly/) package.

```{r}
adsl <- forestly::forestly_adsl
adae <- forestly::forestly_adae

adsl$TRT01A <- factor(
  adsl$TRT01A,
  levels = c("Xanomeline Low Dose", "Placebo"),
  labels = c("Low Dose", "Placebo")
)
adae$TRTA <- factor(
  adae$TRTA,
  levels = c("Xanomeline Low Dose", "Placebo"),
  labels = c("Low Dose", "Placebo")
)

analysis_plan <- metalite::plan(
  analysis = "ae_listing",
  population = "apat",
  observation = "wk12",
  parameter = "ser"
)

meta <- metalite::meta_adam(observation = adae, population = adsl) |>
  metalite::define_plan(analysis_plan) |>
  metalite::define_population(
    name = "apat",
    var = c(
      "USUBJID", "SAFFL", "TRT01A", "TRTDUR",
      "SITEID", "SEX", "RACE", "AGE"
    ),
    group = "TRT01A",
    subset = SAFFL == "Y",
    label = "All Participants as Treated"
  ) |>
  metalite::define_observation(
    name = "wk12",
    var = c(
      "USUBJID", "SAFFL", "TRTA", "AEDECOD", "AEBODSYS", "AEREL",
      "AESER", "AEOUT", "AEACN", "AESDTH", "ASTDT", "AENDT"
    ),
    group = "TRTA",
    subset = SAFFL == "Y",
    label = "Weeks 0 to 12"
  ) |>
  metalite::define_parameter(
    name = "ser",
    term1 = "Serious",
    term2 = "",
    subset = AESER == "Y",
    var = "AEDECOD",
    soc = "AEBODSYS",
    label = "Serious AEs"
  ) |>
  metalite::define_analysis(
    name = "ae_listing",
    var_name = c(
      "USUBJID", "ASTDY", "AEDECOD", "ADURN",
      "AESEV", "AESER", "AEREL", "AEOUT"
    ),
    group_by = c("USUBJID", "ASTDY"),
    page_by = "TRTA"
  ) |>
  metalite::meta_build()
```

## Step 2: Generate the AE listing table

`prepare_ae_listing()` uses the population, observation, parameter, and
analysis definitions in `meta` to prepare the listing dataset. The result is
passed directly to `tlf_ae_listing()` to create the RTF table.

```{r}
footnote <- c(
  "Related: Investigator-assessed relationship of the adverse event to study medication. Y = RELATED, N = NOT RELATED",
  "Action Taken: Discontinued = DRUG WITHDRAWN, Interrupted = DRUG INTERRUPTED, Reduced = DOSE REDUCED, Increased = DOSE INCREASED, None = DOSE NOT CHANGED, N/A = NOT APPLICABLE.",
  "Outcome: Resolved = RECOVERED/RESOLVED, Resolving = RECOVERING/RESOLVING, Sequelae = RECOVERED/RESOLVED WITH SEQUELAE, Not resolved = NOT RECOVERED/NOT RESOLVED.",
  "Adverse event terms are from MedDRA Version 25.0."
)

rtf_file <- tempfile(fileext = ".rtf")

prepare_ae_listing(
  meta,
  analysis = "ae_listing",
  population = "apat",
  observation = "wk12",
  parameter = "ser"
) |>
  tlf_ae_listing(
    footnotes = footnote,
    orientation = "portrait",
    source = "Source:  [CDISCpilot: adam-adsl; adae]",
    analysis = "ae_listing", # Provide analysis type defined in meta$analysis
    path_outtable = rtf_file
  )
```

```{r convert-rtf-to-pdf, include=FALSE}
pdf_dir <- if (dir.exists("vignettes/pdf")) "vignettes/pdf" else "pdf"
pdf_file <- file.path(pdf_dir, "ae0listing1.pdf")

libreoffice <- c(
  Sys.which(c("libreoffice", "soffice")),
  "/Applications/LibreOffice.app/Contents/MacOS/soffice"
)
libreoffice <- unname(libreoffice[file.exists(libreoffice)][1])
browser <- c(
  "/Applications/Google Chrome.app/Contents/MacOS/Google Chrome",
  "/Applications/Microsoft Edge.app/Contents/MacOS/Microsoft Edge"
)
browser <- browser[file.exists(browser)][1]

if (!is.na(libreoffice)) {
  conversion_dir <- tempfile("ae-listing-pdf-")
  dir.create(conversion_dir)

  system2(
    libreoffice,
    c(
      "--headless", "--convert-to", "pdf",
      "--outdir", shQuote(conversion_dir), shQuote(rtf_file)
    ),
    stdout = TRUE,
    stderr = TRUE
  )

  converted_pdf <- file.path(
    conversion_dir,
    paste0(tools::file_path_sans_ext(basename(rtf_file)), ".pdf")
  )

  if (file.exists(converted_pdf)) {
    file.copy(converted_pdf, pdf_file, overwrite = TRUE)
  } else {
    warning("LibreOffice did not create the expected PDF; using the committed PDF.")
  }
} else if (Sys.info()[["sysname"]] == "Darwin" && !is.na(browser)) {
  conversion_dir <- tempfile("ae-listing-pdf-")
  dir.create(conversion_dir)
  html_file <- file.path(conversion_dir, "ae0listing1.html")
  browser_profile <- file.path(conversion_dir, "browser-profile")

  system2(
    "/usr/bin/textutil",
    c("-convert", "html", "-output", shQuote(html_file), shQuote(rtf_file))
  )
  suppressWarnings(
    system2(
      browser,
      c(
        "--headless=new", "--disable-gpu", "--no-pdf-header-footer",
        paste0("--user-data-dir=", shQuote(browser_profile)),
        paste0("--print-to-pdf=", shQuote(normalizePath(pdf_file, mustWork = FALSE))),
        paste0("file://", normalizePath(html_file))
      ),
      stdout = FALSE,
      stderr = FALSE,
      timeout = 10
    )
  )

  if (!file.exists(pdf_file)) {
    stop("The browser did not create the expected PDF.", call. = FALSE)
  }
} else if (!file.exists(pdf_file)) {
  stop(
    paste(
      "No supported RTF-to-PDF converter was found and the PDF does not exist.",
      "Install LibreOffice, then rerun this chunk."
    ),
    call. = FALSE
  )
}
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics(pdf_file)
```
