---
title: "AE Listing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AE Listing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(metalite.ae)
library(metalite)
library(r2rtf)
```

## Overview

The AE listing provides details about the interested adverse events information.
The development of AE listing involves:

- `prepare_ae_listing()`: prepare listing datasets.
- `tlf_ae_listing()`: transfer output dataset to RTF table.

> We will illustrate the process to generate an AE listing RTF file similar to the following one.

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("pdf/ae0listing0ser0wk12.pdf")
```

### Data manipulation

```{r}
adae <- r2rtf::r2rtf_adae
adsl <- r2rtf::r2rtf_adsl
```

First of all, we need to format the data from observation data set as we need to
display on the final output. For example, if we need a 'related' column in the
final output listing, we need to have a 'related' column in the observation
data set.

In the following example, we illustrate how to create a variable 'related'
which use 'Y' and 'N' to present 'RELATED' and 'NOT RELATED' in listing.

```{r}
adae$related <- ifelse(
  adae$AEREL == "RELATED",
  "Y",
  ifelse(
    toupper(adae$AEREL) == "NOT RELATED",
    "N",
    tools::toTitleCase(tolower(adae$AEREL))
  )
)
head(adae$related)
```

In this code chunk, we provide a solution to generate an `outcome` variable
in `adae` used to display AE outcome details in listings.

```{r}
for (i in seq_along(adae$AEOUT)) {
  adae$outcome <- switch(adae$AEOUT[i],
    "RECOVERED/RESOLVED" = "Resolved",
    "RECOVERING/RESOLVING" = "Resolving",
    "RECOVERED/RESOLVED WITH SEQUELAE" = "Sequelae",
    "NOT RECOVERED/NOT RESOLVED" = "Not Resolved",
    tools::toTitleCase(tolower(adae$AEOUT[i]))
  )
}
head(adae$outcome)
```

In this example, we attempt to illustrate how to generate a column for
`Action Taken` information.

```{r}
# Raw data is all blank for 'AEACN',
# change to 'DOSE NOT CHANGED' for better illustration.
adae$AEACN <- gsub("", "DOSE NOT CHANGED", adae$AEACN)

for (i in seq_along(adae$AEACN)) {
  adae$action_taken[i] <- switch(adae$AEACN[i],
    "DOSE NOT CHANGED" = "None",
    "DOSE REDUCED" = "Reduced",
    "DRUG INTERRUPTED" = "Interrupted",
    "DOSE INCREASED" = "Increased",
    "NOT APPLICABLE" = "N/A",
    "UNKNOWN" = "Unknown",
    "''" = "None",
    tools::toTitleCase(tolower(adae$AEACN[i]))
  )
}
head(adae$action_taken)
```

Example to create column for adverse event duration.

```{r}
# AE duration with unit
adae$duration <- paste(
  ifelse(
    is.na(adae$ADURN),
    "",
    as.character(adae$ADURN)
  ),
  tools::toTitleCase(tolower(adae$ADURU)),
  sep = " "
)

for (i in seq_along(adae$duration)) {
  if (is.na(adae$ADURN[i])) {
    adae$duration[i] <- ifelse(
      charmatch(toupper(adae$AEOUT[i]), "RECOVERING/RESOLVING") > 0 |
        charmatch(toupper(adae$AEOUT[i]), "NOT RECOVERED/NOT RESOLVED") > 0,
      "Continuing",
      "Unknown"
    )
  }
}
head(adae$duration)
```

The `subline` variable is used as subline header in the AE listing.
Here, we create the subline variable by concatenating
'Subject ID', 'Gender', 'Race', 'Age', and 'Treatment' information.

```{r}
adae$subline <- paste0(
  "Subject ID = ", adae$USUBJID,
  ", Gender = ", adae$SEX,
  ", Race = ", adae$RACE,
  ", AGE = ", adae$AGE, " Years",
  ", TRT = ", adae$TRTA
)
head(adae$subline)
```

We assign label, which will be the column header of AE listing, to variables used for listing.

```{r, message=FALSE}
adae <- assign_label(adae,
  var = c("related", "outcome", "duration", "AESEV", "AESER", "AEDECOD", "action_taken"),
  label = c("Related", "Outcome", "Duration", "Intensity", "Serious", "Adverse Event", "Action Taken")
)
get_label(adae)
```

### Create `metadata` for AE listing

In order to generate AE listings using metalite.ae, we need to create a
`metadata` object using the metalite package.

While defining `analysis` for listing, several mapping names should be included:

- `name`: The name of this analysis. In this example, we use 'ae_listing'.
- `var_name`: Variable name in observation dataset which used to display
  in the listing. The order in the input vector for `var_nasme` will be
  the order of column display in the final listing.
- `subline_by`: The variable is used as the page subline header.
  It could be `NULL` if not applicable.
- `group_by`: The variable only display when it is first appeared.
  In this example, we define it as `c("USUBJID", "ASTDY")`.
- `page_by`: The variable is used as section header.
  In this example, we define it as `c("TRTA", "subline")`

```{r}
meta_ae_listing <- function() {
  adsl$TRTA <- adsl$TRT01A

  adsl$TRTA <- factor(
    adsl$TRTA,
    levels = c("Placebo", "Xanomeline Low Dose", "Xanomeline High Dose"),
    labels = c("Placebo", "Low Dose", "High Dose")
  )

  adae$TRTA <- factor(
    adae$TRTA,
    levels = c("Placebo", "Xanomeline Low Dose", "Xanomeline High Dose"),
    labels = c("Placebo", "Low Dose", "High Dose")
  )

  plan <- plan(
    analysis = "ae_listing", population = "apat",
    observation = c("wk12", "wk24"), parameter = c("any", "rel", "ser")
  )

  meta_adam(
    population = adsl,
    observation = adae
  ) |>
    define_plan(plan = plan) |>
    define_population(
      name = "apat",
      group = "TRTA",
      subset = quote(SAFFL == "Y")
    ) |>
    define_observation(
      name = "wk12",
      group = "TRTA",
      subset = quote(SAFFL == "Y"),
      label = "Weeks 0 to 12",
      rel_day = "ASTDY"
    ) |>
    define_observation(
      name = "wk24",
      group = "TRTA",
      subset = quote(AOCC01FL == "Y"), # Only for demo, another flag should be used.
      label = "Weeks 0 to 24",
      rel_day = "ASTDY"
    ) |>
    define_parameter(
      name = "rel",
      subset = quote(AEREL %in% c("POSSIBLE", "PROBABLE"))
    ) |>
    define_parameter(
      name = "ser",
      subset = quote(AESER == "Y")
    ) |>
    define_analysis(
      name = "ae_listing",
      var_name = c(
        "USUBJID", "ASTDY", "AEDECOD", "duration",
        "AESEV", "AESER", "related", "action_taken", "outcome"
      ),
      subline_by = NULL,
      group_by = c("USUBJID", "ASTDY"),
      page_by = c("TRTA", "subline")
    ) |>
    meta_build()
}
```

### Analysis preparation

The `prepare_ae_listing()` function is designed to collect the variables used for the listing:

- `meta`: The input is a `meta` object. Here, we use the `meta` object created above.
- `analysis`: Analysis name for listing defined in the input `meta`.
  In this example, it is "ae_listing".

```{r}
meta <- meta_ae_listing()
tbl <- prepare_ae_listing(
  meta,
  analysis = "ae_listing",
  population = "apat",
  observation = "wk12",
  parameter = "ser"
)
head(tbl$tbl)
head(tbl$col_name)
```

### Generate RTF

By using `tlf_ae_listing()`, we can transfer the output from
`prepare_ae_listing()` to an RTF table.

```{r}
footnote <- c(
  "Related: Investigator-assessed relationship of the adverse event to study medication. Y = RELATED, N = NOT RELATED",
  "Action Taken: Discontinued = DRUG WITHDRAWN, Interrupted = DRUG INTERRUPTED, Reduced = DOSE REDUCED, Increased = DOSE INCREASED, None = DOSE NOT CHANGED, N/A = NOT APPLICABLE.",
  "Outcome: Resolved = RECOVERED/RESOLVED, Resolving = RECOVERING/RESOLVING, Sequelae = RECOVERED/RESOLVED WITH SEQUELAE, Not resolved = NOT RECOVERED/NOT RESOLVED.",
  "Adverse event terms are from MedDRA Version 25.0."
)
```

```{r}
tbl |>
  tlf_ae_listing(
    footnotes = footnote,
    orientation = "portrait",
    source = "Source:  [CDISCpilot: adam-adsl; adae]",
    path_outtable = "rtf/ae0listing0ser0wk12.rtf",
    path_outdata = NULL
  )
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("pdf/ae0listing0ser0wk12.pdf")
```
