---
title: "AE Listing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AE Listing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(metalite.ae)
library(metalite)
library(r2rtf)
```


## Overview

The AE listing provides details about the interested adverse events information.
The development of AE listing contains exported functions:

- `prepare_ae_listing`: prepare listing datasets
- `tlf_ae_listting`: transfer output dataset to RTF table

> We will illustrate the process to generate a AE listing RTF file similar to the following one.

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("outtable/ae0listing0ser0wk12.pdf")
```

### Data Manipulation

```{r}
adae <- r2rtf::r2rtf_adae
adsl <- r2rtf::r2rtf_adsl
```

First of all, we need to format the data from observation data set as we need to display on the final output. For example, if we need a 'related' column in the final output listing, we need to have a 'related' column in the observation data set. I

In the following example, we illustrate how to create a variable 'related' which use 'Y' and 'N' to present 'RELATED' and 'NOT RELATED' in listing.

```{r}
adae$related <- ifelse(adae$AEREL == "RELATED", "Y", ifelse(
  toupper(adae$AEREL) == "NOT RELATED", "N", tools::toTitleCase(tolower(adae$AEREL))
))
head(adae$related)
```

In this code chunk, we provide a solution to generate a `outcome` variable in `adae` used to display AE outcome details in listings. 

```{r}
for (i in 1:length(adae$AEOUT)) {
  adae$outcome <- switch(adae$AEOUT[i],
    "RECOVERED/RESOLVED" = "Resolved",
    "RECOVERING/RESOLVING" = "Resolving",
    "RECOVERED/RESOLVED WITH SEQUELAE" = "Sequelae",
    "NOT RECOVERED/NOT RESOLVED" = "Not Resolved",
    tools::toTitleCase(tolower(adae$AEOUT[i]))
  )
}
head(adae$outcome)
```


This example we attempt to illustrate how to generate a column for `Action Taken` information.

```{r}
adae$AEACN <- gsub("", "DOSE NOT CHANGED", adae$AEACN) # raw data is all blank for 'AEACN', change to 'DOSE NOT CHANGED' for better illustration

for (i in 1:length(adae$AEACN)) {
  adae$action_taken[i] <- switch(adae$AEACN[i],
    "DOSE NOT CHANGED" = "None",
    "DOSE REDUCED" = "Reduced",
    "DRUG INTERRUPTED" = "Interrupted",
    "DOSE INCREASED" = "Increased",
    "NOT APPLICABLE" = "N/A",
    "UNKNOWN" = "Unknown",
    "''" = "None",
    tools::toTitleCase(tolower(adae$AEACN[i]))
  )
}
head(adae$action_taken)
```

Example to create column for adverse event duration.

```{r}
adae$duration <- paste(ifelse(is.na(adae$ADURN), "", as.character(adae$ADURN)), tools::toTitleCase(tolower(adae$ADURU)), sep = " ") # AE duration with unit

for (i in 1:length(adae$duration)) {
  if (is.na(adae$ADURN[i])) {
    adae$duration[i] <- ifelse(charmatch(toupper(adae$AEOUT[i]), "RECOVERING/RESOLVING") > 0 |
      charmatch(toupper(adae$AEOUT[i]), "NOT RECOVERED/NOT RESOLVED") > 0, "Continuing", "Unknown")
  }
}
head(adae$duration)
```

`Subline` variable is used as subline header in the AE listing. Here, we create the subline variable by concatenating 'Subject ID', 'Gender', 'Race', 'Age' and 'Treatment' information.

```{r}
adae$subline <- paste0(
  "Subject ID = ", adae$USUBJID,
  ", Gender = ", adae$SEX,
  ", Race = ", adae$RACE,
  ", AGE = ", adae$AGE, " Years",
  ", TRT = ", adae$TRTA
)
head(adae$subline)
```

We assign label, which will be the column header of AE listing, to variables used for listing.

```{r, message=FALSE}
adae <- assign_label(adae,
  var = c("related", "outcome", "duration", "AESEV", "AESER", "AEDECOD", "action_taken"),
  label = c("Related", "Outcome", "Duration", "Intensity", "Serious", "Adverse Event", "Action Taken")
)
get_label(adae)
```

### Create `metadata` for AE listing

In order to generate AE listings using `metalite.ae`, we have to create a `meatadata` using `metalite` package. 

While defining `analysis` for listing, several mapping name should be included:

- `name`: The name of this analysis. In this example, we use 'ae_listing'. 

- `var_name`: Variable name in observation dataset which used to display in the listing. The order in the input vector for `var_nasme` will be the order of column display in the final listing.

- `subline_by`: The variable is used as page subline header. It could be NULL if not applicable.

- `group_by`: The variable only display when it is first appeared. In this example, we define as `c("USUBJID", "ASTDY")`.

- `page_by`: The variable is used as section header. In this example, we define as `c("TRTA", "subline")`

```{r}
meta_ae_listing <- function() {
  adsl$TRTA <- adsl$TRT01A
  adsl$TRTA <- factor(adsl$TRTA,
    levels = c("Placebo", "Xanomeline Low Dose", "Xanomeline High Dose"),
    labels = c("Placebo", "Low Dose", "High Dose")
  )


  adae$TRTA <- factor(adae$TRTA,
    levels = c("Placebo", "Xanomeline Low Dose", "Xanomeline High Dose"),
    labels = c("Placebo", "Low Dose", "High Dose")
  )

  plan <- plan(
    analysis = "ae_listing", population = "apat",
    observation = c("wk12", "wk24"), parameter = c("any", "rel", "ser")
  )

  meta_adam(
    population = adsl,
    observation = adae
  ) |>
    define_plan(plan = plan) |>
    define_population(
      name = "apat",
      group = "TRTA",
      subset = quote(SAFFL == "Y")
    ) |>
    define_observation(
      name = "wk12",
      group = "TRTA",
      subset = quote(SAFFL == "Y"),
      label = "Weeks 0 to 12",
      rel_day = "ASTDY"
    ) |>
    define_observation(
      name = "wk24",
      group = "TRTA",
      subset = quote(AOCC01FL == "Y"), # just for demo, another flag shall be used.
      label = "Weeks 0 to 24",
      rel_day = "ASTDY"
    ) |>
    define_parameter(
      name = "rel",
      subset = quote(AEREL %in% c("POSSIBLE", "PROBABLE"))
    ) |>
    define_parameter(
      name = "ser",
      subset = quote(AESER == "Y")
    ) |>
    define_analysis(
      name = "ae_listing",
      var_name = c("USUBJID", "ASTDY", "AEDECOD", "duration", "AESEV", "AESER", "related", "action_taken", "outcome"),
      subline_by = NULL,
      group_by = c("USUBJID", "ASTDY"),
      page_by = c("TRTA", "subline")
    ) |>
    meta_build()
}
```

### Analysis preparation

The `prepare_ae_listing` function is design to collect the variables used for listing.

- `meta`: The input is a `meta` objects. Here we use the `meta` object we have just created above.

- `analysis`: Analysis name for listing defined in input `meta`. In this example it is "ae_listing".


```{r}
meta <- meta_ae_listing()
tbl <- prepare_ae_listing(meta,
  analysis = "ae_listing",
  population = "apat",
  observation = "wk12",
  parameter = "ser"
)
head(tbl$tbl)
head(tbl$col_name)
```

### Generate RTF

By using `tlf_ae_listing`, we can transfer the output from `prepare_ae_listing` to an RTF table.

```{r}
footnote <- c("Related: Investigator-assessed relationship of the adverse event to study medication. Y = RELATED, N = NOT RELATED",
 "Action Taken: Discontinued = DRUG WITHDRAWN, Interrupted = DRUG INTERRUPTED, Reduced = DOSE REDUCED, Increased = DOSE INCREASED, None = DOSE NOT CHANGED, N/A = NOT APPLICABLE.",
 "Outcome: Resolved = RECOVERED/RESOLVED, Resolving = RECOVERING/RESOLVING, Sequelae = RECOVERED/RESOLVED WITH SEQUELAE, Not resolved = NOT RECOVERED/NOT RESOLVED.",
 "Adverse event terms are from MedDRA Version 25.0.")
```

```{r}
tbl |>
  tlf_ae_listing(
    footnotes = footnote,
    orientation = "portrait",
    source = "Source:  [CDISCpilot: adam-adsl; adae]",
    path_outtable = "outtable/ae0listing0ser0wk12.rtf",
    path_outdata = NULL
  )
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("outtable/ae0listing0ser0wk12.pdf")
```
