---
title: "AE Summary"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AE Summary}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  out.width = "100%",
  dpi = 150,
  eval = TRUE
)
```

```{r}
library(metalite)
library(metalite.ae)
```

## Overview

The purpose of this tutorial is to create production ready AE summary analyses 
by extending examples shown in 
[Chpater 7 of the R for Clinical Study Reports and Submission](https://r4csr.org/aesummary.html) book. 

The AE summary analysis provides tables to summarize adverse events information.
With a metadata created by metalite, 
there are three required functions to create AE summary analysis table using metalite.ae:

- `prepare_ae_summary`: prepare analysis raw datasets
- `format_ae_summary`: prepare analysis (mock) outdata with proper format
- `tlf_ae_summary`: transfer (mock) output dataset to RTF table

An example output: 

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("pdf/ae0summary1.pdf")
```

## Example data 

In metalite.ae, we created an example dataset using ADSL and ADAE datasets by the metalite package. 

```{r}
meta <- meta_ae_example()
meta
```

The same metadata structure is used to support all analyses example in metalite.ae. 
More details can be found in the [metalite package website](https://merck.github.io/metalite/articles/metalite.html). 

## Analysis preparation

The `prepare_ae_summary` function is used to calculate statistics required for 
AE summary analysis using pre-specifed keywords in `meta`.
The input of the function is a `meta` objects created from `metalite` package.
The output of the function is an `outdata` object that contain a list of analysis raw datasets.

```{r, message=FALSE}
outdata <- prepare_ae_summary(meta,
  population = "apat",
  observation = "wk12",
  parameter = "any;rel;ser"
)
```

```{r}
outdata
```

The output dataset contains commonly used statistics. 
The variable is indexed by the order of `outdata$group`.

```{r}
outdata$group
```

The row is indexed by the order of `outdata$name`.

```{r}
head(data.frame(outdata$order, outdata$name))
```

- `n_pop`: participants in population

```{r}
outdata$n_pop
```

- `n`: number of subjects with AE

```{r}
head(outdata$n)
```

- `prop`: proportion of subjects with AE

```{r}
head(outdata$prop)
```

- `diff`: risk difference compared with the `reference_group`.

```{r}
head(outdata$diff)
```

## Format output

After we have the raw analysis results, 
we can use `format_ae_summary` to prepare the outdata
for production ready RTF tables.

```{r}
tbl <- outdata |>
  format_ae_summary()

tbl$tbl
```

### Additional statistics 

The `display` argument allow us to select statistics. For example, we can add risk difference

```{r}
tbl <- outdata |>
  format_ae_summary(display = c("n", "prop", "diff"))

tbl$tbl
```

For advanced analysis, we need `extend_ae_specific_inference`. 
For example, we can add 95% confidence interval based on Miettinen and Nurminen (M&N) method. 
The details of the M&N method can be found in the [rate compare page](https://merck.github.io/metalite.ae/articles/rate-compare.html)

```{r}
tbl <- outdata |>
  extend_ae_specific_inference() |>
  format_ae_summary(display = c("n", "prop", "diff", "diff_ci"))

tbl$tbl
```

### Mock data preparation

The  `mock` argument allow us to create a mock table easily. 

Note: The purpose of the `mock` argument is not to create a comprehensive mock table template,
but a handy way to help user create a mock table that mimics the exact output layout.
Additional work is required to develop a flexible mock table generation tool
(e.g. a dedicated mock table generation package)

```{r}
tbl <- outdata |>
  format_ae_summary(mock = TRUE)

tbl$tbl
```

## RTF tables

The last step is to prepare the RTF table using `tlf_ae_summary`.

```{r}
outdata |>
  format_ae_summary() |>
  tlf_ae_summary(
    source = "Source:  [CDISCpilot: adam-adsl; adae]",
    path_outtable = "rtf/ae0summary1.rtf"
  )
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("pdf/ae0summary1.pdf")
```

The `tlf_ae_summary` function also provide some commonly used argument to customize the table.

```{r}
outdata |>
  format_ae_summary() |>
  tlf_ae_summary(
    source = "Source:  [CDISCpilot: adam-adsl; adae]",
    col_rel_width = c(6, rep(1, 8)),
    text_font_size = 8,
    orientation = "landscape",
    path_outtable = "rtf/ae0summary2.rtf"
  )
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("pdf/ae0summary2.pdf")
```

The mock table can also be generated. 

```{r}
outdata |>
  format_ae_summary(mock = TRUE) |>
  tlf_ae_summary(
    source = "Source:  [CDISCpilot: adam-adsl; adae]",
    path_outtable = "rtf/mock_ae0summary1.rtf"
  )
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("pdf/mock_ae0summary1.pdf")
```
