---
title: "AE Summary"
output:
  rmarkdown::html_vignette
resource_files:
   - package_manual.pdf
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{AE Summary}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  out.width = "100%",
  dpi = 150,
  eval = TRUE
)
```

```{r, include=FALSE}
library(metalite)
library(metalite.ae)
```

## Overview

The AE summary analysis provides tables to summarize adverse events information.
The development of AE summary analysis contains exported functions:

- `prepare_ae_summary`: prepare analysis raw datasets
- `format_ae_summary`: prepare analysis (mock) outdata with proper format
- `tlf_ae_summary`: transfer (mock) output dataset to RTF table

We also prepare (optional) additional function to extend the analysis.

- **TODO** `extend_ae_specific_inference`: add risk difference inference results based on M&N method.

### Analysis preparation

The `prepare_ae_summary` function is design to be used for multiple purposes.
The input of the function is a `meta` objects created from `metalite` package.

```{r}
meta <- meta_ae_dummy()
```

The output of the function is an `outdata` object that contain a list of analysis raw datasets.
The parameter is a group of key words available in the `names(meta$parameter)` and seperated by `;`.

```{r}
outdata <- prepare_ae_summary(meta,
  population = "apat",
  observation = "wk12",
  parameter = "any;rel;ser"
)
```

```{r}
outdata
```

The output dataset contains commonly used statistics. The variable is indexed by the order of `outdata$group`.

```{r}
outdata$group
```

The row is indexed by the order of `outdata$name`.

```{r}
head(data.frame(outdata$order, outdata$name))
```

- `n_pop`: participants in population

```{r}
outdata$n_pop
```

- `n`: number of subjects with AE

```{r}
head(outdata$n)
```

- `prop`: proportion of subjects with AE

```{r}
head(outdata$prop)
```

- `diff`: risk difference compared with the `reference_group`.

```{r}
head(outdata$diff)
```

### Format output

After we have the raw analysis results, we can use `format_ae_summary` to prepare the outdata
to create RTF tables.

```{r}
tbl <- outdata |>
  format_ae_summary()
head(tbl$tbl)
```

We can choose which measurement to display

```{r}
tbl <- outdata |>
  format_ae_specific(display = c("n", "prop"))
head(tbl$tbl)
```

We can add risk difference

```{r}
tbl <- outdata |>
  format_ae_specific(display = c("n", "prop", "diff"))
head(tbl$tbl)
```

We can add 95% confidence interval of M&N method. It requires `extend_ae_specific_inference`
to perform additional calculation.

```{r, eval = FALSE}
tbl <- outdata |>
  extend_ae_specific_inference() |>
  format_ae_specific(display = c("n", "prop", "diff", "diff_ci"))
head(tbl$tbl)
```

### Mock data preparation

We can also use `format_ae_summary` to create mock output data.

The purpose of the `mock` argument is not to create a comprehensive mock table template,
but a handy way to help user create a mock table that mimics the exact output layout.

Additional work is required to develop a flexible mock table generation tool
(e.g. a dedicated mock table generation package)

```{r}
tbl <- outdata |>
  format_ae_summary(mock = TRUE)
head(tbl$tbl)
```

### RTF tables

By using `tlf_ae_summary`, we can transfer the output from `prepare_ae_summary` to an RTF table.

```{r}
outdata |>
  format_ae_summary() |>
  tlf_ae_summary(
    source = "Source:  [CDISCpilot: adam-adsl; adae]",
    path_outtable = "outtable/ae0summary1.rtf"
  )
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("outtable/ae0summary1.pdf")
```

The `tlf_ae_summary` function also provide some commonly used argument to customize the table.

```{r}
outdata |>
  format_ae_summary() |>
  tlf_ae_summary(
    source = "Source:  [CDISCpilot: adam-adsl; adae]",
    col_rel_width = c(6, rep(1, 8)),
    text_font_size = 8,
    orientation = "landscape",
    path_outtable = "outtable/ae0summary2.rtf"
  )
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("outtable/ae0summary2.pdf")
```

We can also generate the mock table

```{r}
outdata |>
  format_ae_summary(mock = TRUE) |>
  tlf_ae_summary(
    source = "Source:  [CDISCpilot: adam-adsl; adae]",
    path_outtable = "outtable/mock_ae0summary1.rtf"
  )
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("outtable/mock_ae0summary1.pdf")
```

## Future developement

### Features

- Enable CI and p-value calculation using M&N method.
- Cross check the capability in SAS standard macro to identify required features.
  - `usermanuals/asr0ae0summary/asr0ae0summary-spec.docx`
  - `usermanuals/asr0ae0summary0analysis/asr0ae0summary0analysis-spec.docx`

### Technical requirements (For future development)

- Avoiding `|>`, using `|>` to reduce dependency of `magrittr`.
- The functions in the `R/` folder should not depend on `dplyr`, `tidyr` etc.
