---
title: "Customize Columns in an AE Summary Table"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Customize Columns to be Displayed in an AE Summary Table}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  out.width = "100%",
  dpi = 150
)
```

```{r}
library(metalite.ae)
```

## Overview

The `display` argument of `format_ae_summary()` controls which statistics
appear in an AE summary table. This vignette demonstrates how to add a risk
difference column to the participant count and percentage columns.

## Define metadata

The example uses ADSL and ADAE data from the
[forestly](https://merck.github.io/forestly/) package.

We define the metadata following the same approach used in the
[AE Summary](ae-summary.html) vignette.

```{r}
adsl <- forestly::forestly_adsl
adae <- forestly::forestly_adae

adsl$TRT01A <- factor(
  adsl$TRT01A,
  levels = c("Xanomeline Low Dose", "Placebo"),
  labels = c("Low Dose", "Placebo")
)
adae$TRTA <- factor(
  adae$TRTA,
  levels = c("Xanomeline Low Dose", "Placebo"),
  labels = c("Low Dose", "Placebo")
)

analysis_plan <- metalite::plan(
  analysis = "ae_summary",
  population = "apat",
  observation = "wk12",
  parameter = "any;rel;ser"
)

meta <- metalite::meta_adam(observation = adae, population = adsl) |>
  metalite::define_plan(analysis_plan) |>
  metalite::define_population(
    name = "apat",
    var = c("USUBJID", "SAFFL", "TRT01A"),
    group = "TRT01A",
    subset = SAFFL == "Y",
    label = "All Participants as Treated"
  ) |>
  metalite::define_observation(
    name = "wk12",
    var = c(
      "USUBJID", "SAFFL", "TRTA", "AEDECOD", "AEBODSYS", "AEREL",
      "AESER"
    ),
    group = "TRTA",
    subset = SAFFL == "Y",
    label = "Weeks 0 to 12"
  ) |>
  metalite::define_parameter(
    name = "any",
    term1 = "",
    term2 = "",
    var = "AEDECOD",
    soc = "AEBODSYS",
    label = "All AEs"
  ) |>
  metalite::define_parameter(
    name = "rel",
    term1 = "Drug-Related",
    term2 = "",
    subset = AEREL %in% c("POSSIBLE", "PROBABLE"),
    var = "AEDECOD",
    soc = "AEBODSYS",
    label = "Drug-related AEs"
  ) |>
  metalite::define_parameter(
    name = "ser",
    term1 = "Serious",
    term2 = "",
    subset = AESER == "Y",
    var = "AEDECOD",
    soc = "AEBODSYS",
    label = "Serious AEs"
  ) |>
  metalite::define_analysis(
    name = "ae_summary",
    title = "Adverse Event Summary"
  ) |>
  metalite::meta_build()
```

## Select columns

To perform advanced analysis, the `extend_ae_specific_inference()`
function is utilized.
For instance, we can incorporate a 95% confidence interval based on the
Miettinen and Nurminen (M&N) method.
Further information regarding the M&N method can be found in the
[rate compare vignette](https://merck.github.io/metalite.ae/articles/rate-compare.html).

After the usage of `extend_ae_specific_inference()`, 
use `display = ...` in the `format_ae_summary()` function to choose specific
statistics and set their order in the table. Available options are:

- `"n"`: number of participants with an adverse event.
- `"prop"`: proportion of participants with an adverse event.
- `"total"`: total columns.
- `"diff"`: risk difference.
- `"diff_ci"`: 95% confidence interval for the risk difference using the
  Miettinen and Nurminen method.
- `"diff_p"`: p-value for the risk difference using the Miettinen and
  Nurminen method.
- `"dur"`: average adverse event duration.
- `"events_avg"`: average number of adverse events per participant.
- `"events_count"`: number of adverse events per participant.

The `"diff_ci"` and `"diff_p"` statistics are added by
`extend_ae_specific_inference()`, `"dur"` is added by
`extend_ae_specific_duration()`, and the event statistics are added by
`extend_ae_specific_events()`.

For example, include `"diff"` in addition to the number and proportion of
participants with an adverse event:

```{r}
prepare_ae_summary(
  meta,
  population = "apat",
  observation = "wk12",
  parameter = "any;rel;ser"
) |>
  extend_ae_specific_inference() |>
  format_ae_summary(display = c("n", "prop", "diff", "diff_ci")) |>
  tlf_ae_summary(
    source = "Source:  [CDISCpilot: adam-adsl; adae]",
    analysis = "ae_summary", # Provide analysis type defined in meta$analysis
    col_rel_width = c(3, rep(1, 6)),
    path_outtable = "rtf/ae0summary2.rtf"
  )
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("pdf/ae0summary2.pdf")
```
