---
title: "AE Summary"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AE Summary}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  out.width = "100%",
  dpi = 150,
  eval = TRUE
)
```

```{r, include=FALSE}
library(metalite)
library(metalite.ae)
```

## Overview

The AE summary analysis provides tables to summarize adverse events information.
There are three steps to create an AE summary analysis table using metalite.ae:

- `prepare_ae_summary()`: prepare analysis raw datasets.
- `format_ae_summary()`: prepare analysis (mock) outdata with proper format.
- `tlf_ae_summary()`: transfer (mock) output dataset to RTF table.

With the three functions, we can quickly create an AE summary analysis as shown
in the [AE summary chapter](https://r4csr.org/aesummary.html) of the
_R for Clinical Study Reports and Submission_ book.

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("pdf/ae0summary1.pdf")
```

### Analysis preparation

The `prepare_ae_summary()` function is designed to be used for multiple purposes.
The input of the function is a `meta` object created by the metalite package.

```{r}
meta <- meta_ae_example()
```

The output of the function is an `outdata` object that contains a list of analysis raw datasets.
The `parameter` is a group of key words available in the `names(meta$parameter)` and separated by `;`.

```{r}
outdata <- prepare_ae_summary(meta,
  population = "apat",
  observation = "wk12",
  parameter = "any;rel;ser"
)
```

```{r}
outdata
```

The output dataset contains commonly used statistics. The variable is indexed by the order of `outdata$group`.

```{r}
outdata$group
```

The row is indexed by the order of `outdata$name`.

```{r}
head(data.frame(outdata$order, outdata$name))
```

- `n_pop`: participants in population.

```{r}
outdata$n_pop
```

- `n`: number of subjects with AE.

```{r}
head(outdata$n)
```

- `prop`: proportion of subjects with AE.

```{r}
head(outdata$prop)
```

- `diff`: risk difference compared with the `reference_group`.

```{r}
head(outdata$diff)
```

### Format output

After we have the raw analysis results, we can use `format_ae_summary()` to prepare the outdata
to create RTF tables.

```{r}
tbl <- outdata |>
  format_ae_summary()
head(tbl$tbl)
```

We can choose which measurement to display

```{r}
tbl <- outdata |>
  format_ae_summary(display = c("n", "prop"))
head(tbl$tbl)
```

We can add risk difference

```{r}
tbl <- outdata |>
  format_ae_summary(display = c("n", "prop", "diff"))
head(tbl$tbl)
```

We can add 95% confidence interval of M&N method. It requires `extend_ae_specific_inference()`
to perform additional calculations.

```{r, eval = FALSE}
tbl <- outdata |>
  extend_ae_specific_inference() |>
  format_ae_summary(display = c("n", "prop", "diff", "diff_ci"))
head(tbl$tbl)
```

### Mock data preparation

We can also use `format_ae_summary()` to create mock output data.

The purpose of the `mock` argument is not to create a comprehensive mock table template,
but a handy way to help users create a mock table that mimics the exact output layout.

Additional work is required to develop a flexible mock table generation tool
(for example, a dedicated mock table generation package).

```{r}
tbl <- outdata |>
  format_ae_summary(mock = TRUE)
head(tbl$tbl)
```

### RTF tables

By using `tlf_ae_summary()`, we can transfer the output from `prepare_ae_summary()` to an RTF table.

```{r}
outdata |>
  format_ae_summary() |>
  tlf_ae_summary(
    source = "Source:  [CDISCpilot: adam-adsl; adae]",
    path_outtable = "rtf/ae0summary1.rtf"
  )
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("pdf/ae0summary1.pdf")
```

The `tlf_ae_summary()` function also provides some commonly used argument to customize the table.

```{r}
outdata |>
  format_ae_summary() |>
  tlf_ae_summary(
    source = "Source:  [CDISCpilot: adam-adsl; adae]",
    col_rel_width = c(6, rep(1, 8)),
    text_font_size = 8,
    orientation = "landscape",
    path_outtable = "rtf/ae0summary2.rtf"
  )
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("pdf/ae0summary2.pdf")
```

We can also generate the mock table

```{r}
outdata |>
  format_ae_summary(mock = TRUE) |>
  tlf_ae_summary(
    source = "Source:  [CDISCpilot: adam-adsl; adae]",
    path_outtable = "rtf/mock_ae0summary1.rtf"
  )
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("pdf/mock_ae0summary1.pdf")
```
