---
title: "Filter and Sort an AE Specific Table"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Filter and Sort an AE Specific Table}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  out.width = "100%",
  dpi = 150
)
```

```{r}
library(metalite.ae)
```

## Overview

AE specific tables can contain many system organ classes and preferred terms.
The filtering and sorting arguments of `format_ae_specific()` help focus the
output on clinically relevant rows and present them in a useful order.

## Define metadata

The example uses ADSL and ADAE data from the
[forestly](https://merck.github.io/forestly/) package. The metadata follows the
same approach used in the [AE Specific](ae-specific.html) vignette.

```{r}
adsl <- forestly::forestly_adsl
adae <- forestly::forestly_adae

adsl$TRT01A <- factor(
  adsl$TRT01A,
  levels = c("Xanomeline Low Dose", "Placebo"),
  labels = c("Low Dose", "Placebo")
)
adae$TRTA <- factor(
  adae$TRTA,
  levels = c("Xanomeline Low Dose", "Placebo"),
  labels = c("Low Dose", "Placebo")
)

analysis_plan <- metalite::plan(
  analysis = "ae_specific",
  population = "apat",
  observation = "wk12",
  parameter = "rel"
)

meta <- metalite::meta_adam(observation = adae, population = adsl) |>
  metalite::define_plan(analysis_plan) |>
  metalite::define_population(
    name = "apat",
    var = c(
      "USUBJID", "SAFFL", "TRT01A", "TRTDUR",
      "SITEID", "SEX", "RACE", "AGE"
    ),
    group = "TRT01A",
    subset = SAFFL == "Y",
    label = "All Participants as Treated"
  ) |>
  metalite::define_observation(
    name = "wk12",
    var = c(
      "USUBJID", "SAFFL", "TRTA", "AEDECOD", "AEBODSYS", "AEREL",
      "AESER", "AEOUT", "AEACN", "AESDTH", "ASTDT", "AENDT"
    ),
    group = "TRTA",
    subset = SAFFL == "Y",
    label = "Weeks 0 to 12"
  ) |>
  metalite::define_parameter(
    name = "rel",
    term1 = "Drug-Related",
    term2 = "",
    subset = AEREL %in% c("POSSIBLE", "PROBABLE"),
    var = "AEDECOD",
    soc = "AEBODSYS",
    label = "Drug-related AEs"
  ) |>
  metalite::define_analysis(
    name = "ae_specific",
    title = "Participants with Drug-Related Adverse Events"
  ) |>
  metalite::meta_build()
```

## Filter rows

Set `filter_method` to `"percent"` or `"count"`, then use `filter_criteria`
to define the minimum incidence required in at least one treatment group.
Percentage criteria must be between 0 and 100; count criteria must be greater
than 0.

The following example retains rows where at least one treatment group has an
incidence of 6% or greater. 
To filter by participant count instead, set `filter_method = "count"` and pass
the minimum count to `filter_criteria`.

```{r}
prepare_ae_specific(
  meta,
  population = "apat",
  observation = "wk12",
  parameter = "rel"
) |>
  format_ae_specific(
    filter_method = "percent",
    filter_criteria = 6
  ) |>
  tlf_ae_specific(
    source = "Source:  [CDISCpilot: adam-adsl; adae]",
    analysis = "ae_specific",
    meddra_version = "24.0",
    path_outtable = "rtf/ae0specific4.rtf"
  )
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("pdf/ae0specific4.pdf")
```



## Sort rows

The `sort_order` argument accepts:

- `"alphabetical"`: sort preferred terms alphabetically.
- `"count_des"`: sort participant counts in descending order.
- `"count_asc"`: sort participant counts in ascending order.

For count-based sorting, `sort_column` identifies the treatment group whose
counts determine the order. Its value must match an entry in `outdata$group`.

The following example sorts rows by the Placebo participant count in descending
order:

```{r}
prepare_ae_specific(
  meta,
  population = "apat",
  observation = "wk12",
  parameter = "rel"
) |>
  format_ae_specific(
    sort_order = "count_des",
    sort_column = "Placebo"
  ) |>
  tlf_ae_specific(
    source = "Source:  [CDISCpilot: adam-adsl; adae]",
    analysis = "ae_specific",
    meddra_version = "24.0",
    path_outtable = "rtf/ae0specific5.rtf"
  )
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("pdf/ae0specific5.pdf")
```

## Filter and sort rows

Filtering and sorting can be combined in one call. Filtering is applied first,
and the retained rows are then sorted using the requested treatment group.

```{r}
prepare_ae_specific(
  meta,
  population = "apat",
  observation = "wk12",
  parameter = "rel"
) |>
  format_ae_specific(
    filter_method = "percent",
    filter_criteria = 6,
    sort_order = "count_des",
    sort_column = "Placebo"
  ) |>
  tlf_ae_specific(
    source = "Source:  [CDISCpilot: adam-adsl; adae]",
    analysis = "ae_specific",
    meddra_version = "24.0",
    path_outtable = "rtf/ae0specific6.rtf"
  )
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("pdf/ae0specific6.pdf")
```
