---
title: "AE Specification"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AE Specification}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  out.width = "100%",
  dpi = 150
)
```

```{r}
library(metalite)
library(metalite.ae)
```

## Overview

The purpose of this tutorial is to create production ready AE specification analyses
by extending examples shown in the
[specific AE chapter](https://r4csr.org/specific-ae.html)
of the _R for Clinical Study Reports and Submission_ book.

The AE specification analysis is to provide tables to summarize details of different types of adverse events.
With a metadata object created by metalite,
there are three required functions to create AE specification analysis tables using metalite.ae:

- `prepare_ae_specific()`: prepare analysis raw datasets.
- `format_ae_specific()`: prepare analysis (mock) outdata with proper format.
- `tlf_ae_specific()`: transfer (mock) output dataset to RTF table.

There are three optional functions to extend AE specification analysis.

- `extend_ae_specific_inference()`: add risk difference inference results based on M&N method.
- `extend_ae_specific_duration()`: add average duration of AE.
- `extend_ae_specific_events()`: add average number of AE events.

## Example data

In metalite.ae, we created an example dataset using ADSL and ADAE datasets from the metalite package.

```{r}
meta <- meta_ae_example()
meta
```

The same metadata structure is used to support all analysis examples in metalite.ae.
More details can be found on the [metalite package website](https://merck.github.io/metalite/articles/metalite.html).

### Analysis preparation

The `prepare_ae_specific()` function is used to calculate statistics required for
AE specification analysis using pre-specified keywords in `meta`.
The input of the function is a `meta` object created by metalite.
The output of the function is an `outdata` object that contains a list of analysis raw datasets.

```{r, message = FALSE}
outdata <- prepare_ae_specific(
  meta,
  population = "apat",
  observation = "wk12",
  parameter = "rel"
)
```

```{r}
outdata
```

The output dataset contains commonly used statistics.
The variable is indexed by the order of `outdata$group`.

```{r}
outdata$group
```

The row is indexed by the order of `outdata$name`.

```{r}
head(data.frame(outdata$order, outdata$name))
```

- `n_pop`: participants in population.

```{r}
outdata$n_pop
```

- `n`: number of subjects with AE.

```{r}
head(outdata$n)
```

- `prop`: proportion of subjects with AE.

```{r}
head(outdata$prop)
```

- `diff`: risk difference compared with the `reference_group`.

```{r}
head(outdata$diff)
```

## Format output

After we have the raw analysis results,
we can use `format_ae_specific()` to prepare the outdata
for production ready RTF tables.

```{r}
tbl <- outdata |> format_ae_specific()
head(tbl$tbl)
```

### Additional statistics

The `display` argument allows us to select statistics. For example, we can add risk difference.

```{r}
tbl <- outdata |> format_ae_specific(display = c("n", "prop", "diff"))
head(tbl$tbl)
```

For advanced analysis, we need `extend_*()` functions.
For example, we can use `extend_ae_specific_inference()` to add
95% confidence interval based on Miettinen and Nurminen (M&N) method.
The details of the M&N method can be found in the
[rate compare vignette](https://merck.github.io/metalite.ae/articles/rate-compare.html)

```{r}
tbl <- outdata |>
  extend_ae_specific_inference() |>
  format_ae_specific(display = c("n", "prop", "diff", "diff_ci"))
head(tbl$tbl)
```

We can use `extend_ae_specific_duration()` to add average duration of AE.

```{r}
tbl <- outdata |>
  extend_ae_specific_duration(duration_var = "ADURN") |>
  format_ae_specific(display = c("n", "prop", "dur"))

head(tbl$tbl)
```

We can use `extend_ae_specific_events()` to add average number of AE per subject.

```{r}
tbl <- outdata |>
  extend_ae_specific_events() |>
  format_ae_specific(display = c("n", "prop", "events"))

head(tbl$tbl)
```

### Mock data preparation

The  `mock` argument allows us to create a mock table easily.

Note: The purpose of the `mock` argument is not to create a comprehensive mock table template,
but a handy way to help user create a mock table that mimics the exact output layout.

Additional work is required to develop a flexible mock table generation tool
(for example, a dedicated mock table generation package).

```{r}
tbl <- outdata |> format_ae_specific(mock = TRUE)
head(tbl$tbl)
```

## RTF tables

The last step is to prepare the RTF table using `tlf_ae_specific()`.

```{r}
outdata |>
  format_ae_specific() |>
  tlf_ae_specific(
    meddra_version = "24.0",
    source = "Source:  [CDISCpilot: adam-adsl; adae]",
    path_outtable = "rtf/ae0specific1.rtf"
  )
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("pdf/ae0specific1.pdf")
```

The `tlf_ae_specific()` function also provides some commonly used arguments to customize the table.

```{r}
outdata |>
  format_ae_specific() |>
  tlf_ae_specific(
    meddra_version = "24.0",
    source = "Source:  [CDISCpilot: adam-adsl; adae]",
    col_rel_width = c(6, rep(1, 8)),
    text_font_size = 8,
    orientation = "landscape",
    path_outtable = "rtf/ae0specific2.rtf"
  )
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("pdf/ae0specific2.pdf")
```

The mock table can also be generated.

```{r}
outdata |>
  format_ae_specific(mock = TRUE) |>
  tlf_ae_specific(
    meddra_version = "24.0",
    source = "Source:  [CDISCpilot: adam-adsl; adae]",
    path_outtable = "rtf/mock_ae0specific1.rtf"
  )
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("pdf/mock_ae0specific1.pdf")
```
