---
title: "Generate a Static AE Specific Table"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generate a Static AE Specific Table}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  out.width = "100%",
  dpi = 150
)
```

```{r}
library(metalite.ae)
```

## Overview

This vignette demonstrates how to generate a static AE specific table.

The workflow uses three functions from
[metalite.ae](https://merck.github.io/metalite.ae/):

- `prepare_ae_specific()` prepares the analysis datasets.
- `format_ae_specific()` formats the results for reporting.
- `tlf_ae_specific()` creates the RTF table.

Additional vignettes demonstrate how to extend the displayed statistics,
filter and sort rows, customize the RTF layout, and create mock output.

An example output is shown below.

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("pdf/ae0specific1.pdf")
```

## Procedure to generate an AE specific table

The example uses ADSL and ADAE data from the
[forestly](https://merck.github.io/forestly/) package.

### Step 1: Define metadata

```{r}
# Define metadata
adsl <- forestly::forestly_adsl
adae <- forestly::forestly_adae

adsl$TRT01A <- factor(
  adsl$TRT01A,
  levels = c("Xanomeline Low Dose", "Placebo"),
  labels = c("Low Dose", "Placebo")
)
adae$TRTA <- factor(
  adae$TRTA,
  levels = c("Xanomeline Low Dose", "Placebo"),
  labels = c("Low Dose", "Placebo")
)

analysis_plan <- metalite::plan(
  analysis = "ae_specific",
  population = "apat",
  observation = "wk12",
  parameter = "rel"
)

meta <- metalite::meta_adam(observation = adae, population = adsl) |>
  metalite::define_plan(analysis_plan) |>
  metalite::define_population(
    name = "apat",
    var = c(
      "USUBJID", "SAFFL", "TRT01A", "TRTDUR",
      "SITEID", "SEX", "RACE", "AGE"
    ),
    group = "TRT01A",
    subset = SAFFL == "Y",
    label = "All Participants as Treated"
  ) |>
  metalite::define_observation(
    name = "wk12",
    var = c(
      "USUBJID", "SAFFL", "TRTA", "AEDECOD", "AEBODSYS", "AEREL",
      "AESER", "AEOUT", "AEACN", "AESDTH", "ASTDT", "AENDT"
    ),
    group = "TRTA",
    subset = SAFFL == "Y",
    label = "Weeks 0 to 12"
  ) |>
  metalite::define_parameter(
    name = "rel",
    term1 = "Drug-Related",
    term2 = "",
    subset = AEREL %in% c("POSSIBLE", "PROBABLE"),
    var = "AEDECOD",
    soc = "AEBODSYS",
    label = "Drug-related AEs"
  ) |>
  metalite::define_analysis(
    name = "ae_specific",
    title = "Patients with Drug-Related Adverse Events"
  ) |>
  metalite::meta_build()
```

<details>
<summary>Click to show the output</summary>
```{r}
meta
```
</details>

### Analysis preparation

The function  `prepare_ae_specific()` is used to create a dataset for
AE summary analysis by utilizing predefined keywords specified
in the example data `meta`.

The resulting output of the function is an `outdata` object, which comprises
a collection of raw datasets for analysis and reporting.

```{r, message = FALSE}
outdata <- prepare_ae_specific(
  meta,
  population = "apat",
  observation = "wk12",
  parameter = "rel"
)
```

```{r}
outdata
```

The resulting dataset contains frequently used statistics,
with variables indexed according to the order specified in `outdata$group`.

```{r}
outdata$group
```

The row is indexed according to the order of `outdata$name`.

```{r}
head(data.frame(outdata$order, outdata$name))
```

- `n_pop`: number of participants in population.

```{r}
outdata$n_pop
```

- `n`: number of subjects with AE.

```{r}
head(outdata$n)
```

- `prop`: proportion of subjects with AE.

```{r}
head(outdata$prop)
```

- `diff`: risk difference compared with the `reference_group`.

```{r}
head(outdata$diff)
```

## Format output

Once the raw analysis results are obtained,
the `format_ae_specific()` function can be employed to prepare the outdata,
ensuring its compatibility with production-ready RTF tables.

```{r}
tbl <- outdata |> format_ae_specific()
head(tbl$tbl)
```

### Additional statistics

By using the `display` argument,
we can choose specific statistics to include.
For instance, we have the option to incorporate the risk difference.

```{r}
tbl <- outdata |> format_ae_specific(display = c("n", "prop", "diff"))
head(tbl$tbl)
```

To perform advanced analysis, the `extend_ae_specific_inference()` function
is utilized.
For instance, we can incorporate a 95% confidence interval based on the
Miettinen and Nurminen (M&N) method.
Further information regarding the M&N method can be found in the
[rate compare vignette](https://merck.github.io/metalite.ae/articles/rate-compare.html).

```{r}
tbl <- outdata |>
  extend_ae_specific_inference() |>
  format_ae_specific(display = c("n", "prop", "diff", "diff_ci"))
head(tbl$tbl)
```

We can use `extend_ae_specific_duration()` to add average duration of AE.

```{r}
tbl <- outdata |>
  extend_ae_specific_duration(duration_var = "ADURN") |>
  format_ae_specific(display = c("n", "prop", "dur"))

head(tbl$tbl)
```

We can use `extend_ae_specific_events()` to add number of AE and/or average of it per subject.

```{r}
tbl <- outdata |>
  extend_ae_specific_events() |>
  format_ae_specific(display = c("n", "prop", "events_count", "events_avg"))

head(tbl$tbl)
```

We can use `filter_method` and `filter_criteria` parameters to filter information based on the specified criteria:

- `filter_method`: A character value to specify how to filter rows (by `count` or `percent`).
    - `count`: Filter based on participant count.
    - `percent`: Filter based on percent incidence.
- `filter_criteria`: A numeric value to display rows where at least one therapy group has:
    - a percent incidence or participant count greater than or equal to the specified value.
    - If `filter_method` is `percent`, the value should be between 0 and 100.
    - If `filter_method` is `count`, the value should be greater than 0.


```{r}
tbl <- outdata |>
  extend_ae_specific_events() |>
  format_ae_specific(
    display = c("n", "prop", "events_count", "events_avg"),
    filter_method = "percent",
    filter_criteria = 6
  )

head(tbl$tbl)
```
In results above, rows having any one of "prop_x" values are greater than 6 get kept in the output.


We can use `sort_order` and `sort_column` parameters to sort results based on the specified criteria:

- `sort_order` A character value to specify sorting order:
    - `alphabetical`: Sort by alphabetical order.
    - `count_des`: Sort by count in descending order.
    - `count_asc`: Sort by count in ascending order.
- `sort_column A` character value of `group` in `outdata` used to sort a table with.

```{r}
tbl <- outdata |>
  extend_ae_specific_events() |>
  format_ae_specific(
    display = c("n", "prop", "events_count", "events_avg"),
    sort_order = c("count_des"),
    sort_column = c("Placebo")
  )

head(tbl$tbl)
```

### Mock data preparation

The `mock` argument facilitates the creation of a mock table with ease.

Please note that the intention of the `mock` argument is not to provide
an all-encompassing mock table template.
Instead, it serves as a convenient method to assist users in generating
a mock table that closely resembles the desired output layout.
To develop a more versatile mock table generation tool, further efforts
are necessary.
This could potentially involve the creation of a dedicated mock table
generation package or similar solutions.

```{r}
tbl <- outdata |> format_ae_specific(mock = TRUE)
head(tbl$tbl)
```

## RTF tables

The last step is to prepare the RTF table using `tlf_ae_summary()`.

```{r}
outdata |>
  format_ae_specific() |>
  tlf_ae_specific(
    meddra_version = "24.0",
    source = "Source:  [CDISCpilot: adam-adsl; adae]",
    analysis = "ae_specific", # Provide analysis type defined in meta$analysis
    path_outtable = "rtf/ae0specific1.rtf"
  )
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("pdf/ae0specific1.pdf")
```

The `tlf_ae_specific()` function also provides some commonly used arguments
to customize the table.

```{r}
outdata |>
  format_ae_specific() |>
  tlf_ae_specific(
    meddra_version = "24.0",
    source = "Source:  [CDISCpilot: adam-adsl; adae]",
    analysis = "ae_specific", # Provide analysis type defined in meta$analysis
    col_rel_width = c(6, rep(1, 6)),
    text_font_size = 8,
    orientation = "landscape",
    path_outtable = "rtf/ae0specific2.rtf"
  )
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("pdf/ae0specific2.pdf")
```

The mock table can also be generated.

```{r}
outdata |>
  format_ae_specific(mock = TRUE) |>
  tlf_ae_specific(
    meddra_version = "24.0",
    source = "Source:  [CDISCpilot: adam-adsl; adae]",
    analysis = "ae_specific", # Provide analysis type defined in meta$analysis
    path_outtable = "rtf/mock_ae0specific1.rtf"
  )
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("pdf/mock_ae0specific1.pdf")
```
