---
title: "AE Specification for subgroup analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AE Specification for subgroup analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  out.width = "100%",
  dpi = 150
)
```

```{r, include=FALSE}
library(metalite)
# library(metalite.ae)
devtools::load_all()
```

```{r, include=FALSE}
library(dplyr)
library(tidyr)
```

## Overview

The AE specific Subgroup analysis is to provide tables to summarize details of adverse events by Subgroup.
The development of AE specific Subgroup analysis contain exported functions

- `prepare_ae_specific_subgroup`: prepare analysis raw datasets
- `format_ae_specific_subgroup`: prepare analysis (mock) outdata with proper format
- `tlf_ae_specific_subgroup`: transfer (mock) output dataset to RTF table

### Analysis preparation

The `prepare_ae_specific_subgroup` function is design to be used for multiple purpose.
The input of the function is a `meta` objects created from `metalite.ae` package.

```{r}
meta <- meta_ae_example()
```

The output of the function is an `outdata` objects contain a list of analysis raw datasets.
Key information needed are `subgroup_var`, `display_subgroup_total` and

```{r}
outdata <- prepare_ae_specific_subgroup(meta,
                               population = "apat",
                               observation = "wk12",
                               parameter = "rel",
                               subgroup_var = "SEX",
                               subgroup_header = c("SEX", "TRTA"),
                               display_subgroup_total = T)
```

```{r}
outdata
```

The output dataset contains commonly used statistics within each `subgroup_var`.

```{r}
outdata$out_all$`AMERICAN INDIAN OR ALASKA NATIVE`
outdata$out_all$`BLACK OR AFRICAN AMERICAN`
outdata$out_all$WHITE
outdata$out_all$Total

```

The variable is indexed by the order of `outdata$group` and `outdata$subgroup` within each `subgroup_var`.

```{r}
outdata$group
outdata$subgroup
```

The row is indexed by the order of `name` within each `subgroup_var` analysis output.

```{r}
head(data.frame(outdata$out_all$Total$order, outdata$out_all$Total$name))
```

- `n_pop`: participants in population within each `subgroup_var`.

```{r}
outdata$out_all$`AMERICAN INDIAN OR ALASKA NATIVE`$n_pop
outdata$out_all$`BLACK OR AFRICAN AMERICAN`$n_pop
outdata$out_all$WHITE$n_pop
outdata$out_all$Total$n_pop
```

- `n`: number of subjects with AE whitin each `subgroup_var`.

```{r}
head(outdata$out_all$`AMERICAN INDIAN OR ALASKA NATIVE`$n)
head(outdata$out_all$`BLACK OR AFRICAN AMERICAN`$n)
head(outdata$out_all$WHITE$n)
head(outdata$out_all$Total$n)
```

- `prop`: proportion of subjects with AE within each `subgroup_var`.

```{r}
head(outdata$out_all$`AMERICAN INDIAN OR ALASKA NATIVE`$prop)
head(outdata$out_all$`BLACK OR AFRICAN AMERICAN`$prop)
head(outdata$out_all$WHITE$prop)
head(outdata$out_all$Total$prop)
```

- `diff`: risk difference compared with the `reference_group` within each `subgroup_var`.

```{r}
head(outdata$out_all$Total$diff)
```

### Format output

After we have the raw analysis results, we can use `format_ae_specific_subgroup` to prepare the outdata
to create RTF tables.

```{r}
tbl <- outdata |>
  format_ae_specific_subgroup()
head(tbl$tbl)
```

We can hide total column

```{r}
tbl <- outdata |>
  format_ae_specific_subgroup(display = c("n", "prop"))
head(tbl$tbl)
```

Adding risk difference

```{r}
tbl <- outdata |>
  format_ae_specific_subgroup(display = c("n", "prop", "diff"))
head(tbl$tbl)
```


### Mock data preparation

We can also use `format_ae_specific_subgroup` to create mock output data.

The purpose of the `mock` argument is not to create a comprehensive mock table template,
but a handy way to help user create a mock table that mimic the exact output layout.

Additional work is required to develop a flexible mock table generation tool
(e.g. a dedicated mock table generation package)

```{r}
tbl <- outdata |>
  format_ae_specific_subgroup(mock = TRUE)
head(tbl$tbl)
```

### RTF tables

By using `tlf_ae_specific`, we can transfer the output from `format_ae_specific` to an RTF table.

```{r}
outdata |>
  format_ae_specific_subgroup() |>
  tlf_ae_specific_subgroup(
    meddra_version = "24.0",
    source = "Source:  [CDISCpilot: adam-adsl; adae]",
    path_outtable = "rtf/ae0specific0sub0race2.rtf"
  )
```

```{r}
outdata |>
  format_ae_specific_subgroup() |>
  tlf_ae_specific_subgroup(
    medra_version = "24.0",
    source = "Source:  [CDISCpilot: adam-adsl; adae]",
    path_outtable = "rtf/ae0specific0sub0race1.rtf"
  )
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("outtable/ae0specific0sub0race1.pdf")
```

The `tlf_ae_specific` function also provide some commonly used argument to customize the table.

```{r}
outdata |>
  format_ae_specific() |>
  tlf_ae_specific(
    medra_version = "24.0",
    source = "Source:  [CDISCpilot: adam-adsl; adae]",
    col_rel_width = c(6, rep(1, 8)),
    text_font_size = 8,
    orientation = "landscape",
    path_outtable = "outtable/ae0specific2.rtf"
  )
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("outtable/ae0specific2.pdf")
```

We can also generate the mock table

```{r}
outdata |>
  format_ae_specific(mock = TRUE) |>
  tlf_ae_specific(
    medra_version = "24.0",
    source = "Source:  [CDISCpilot: adam-adsl; adae]",
    path_outtable = "outtable/mock_ae0specific1.rtf"
  )
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("outtable/mock_ae0specific1.pdf")
```

## Future developement

### Features

- Cross check the capability in SAS standard macro to identify required features.

### Technical requirements (for future development)

- Avoiding `%>%`, using `|>` to reduce dependency of `magrittr`.
- The functions in the `R/` folder should not depends on `dplyr`, `tidyr` etc.
