---
title: "AE Specification"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AE Specification}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
resource_files:
   - package_manual.pdf
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  out.width = "100%",
  dpi = 150
)
```

```{r, include=FALSE}
library(metalite)
library(metalite.ae)
```

```{r, include=FALSE}
library(dplyr)
library(tidyr)
```

## Overview

The AE specification analysis is to provide tables to summarize details of adverse events.
The development of AE specification analysis contain exported functions

- `prepare_ae_specific`: prepare analysis raw datasets
- `format_ae_specific`: prepare analysis (mock) outdata with proper format
- `tlf_ae_specific`: transfer (mock) output dataset to RTF table

We also prepare (optional) additional function to extend the analysis.

- `extend_ae_specific_inference`: add risk difference inference results based on M&N method.
- `extend_ae_specific_duration`: add average duration of AE.
- `extend_ae_specific_events`: add average number of AE events.

### Analysis preparation

The `prepare_ae_specific` function is design to be used for multiple purpose.
The input of the function is a `meta` objects created from `metalite.ae` package.

```{r}
meta <- meta_ae_example()
```

The output of the function is an `outdata` objects contain a list of analysis raw datasets.

```{r}
outdata <- prepare_ae_specific(meta,
  population = "apat",
  observation = "wk12",
  parameter = "rel"
)
```

```{r}
outdata
```

The output dataset contains commonly used statistics. The variable is indexed by the order of `outdata$group`.

```{r}
outdata$group
```

The row is indexed by the order of `outdata$name`.

```{r}
head(data.frame(outdata$order, outdata$name))
```

- `n_pop`: participants in population

```{r}
outdata$n_pop
```

- `n`: number of subjects with AE

```{r}
head(outdata$n)
```

- `prop`: proportion of subjects with AE

```{r}
head(outdata$prop)
```

- `diff`: risk difference compared with the `reference_group`.

```{r}
head(outdata$diff)
```

### Format output

After we have the raw analysis results, we can use `format_ae_specific` to prepare the outdata
to create RTF tables.

```{r}
tbl <- outdata |>
  format_ae_specific()
head(tbl$tbl)
```

We can hide total column

```{r}
tbl <- outdata |>
  format_ae_specific(display = c("n", "prop"))
head(tbl$tbl)
```

Adding risk difference

```{r}
tbl <- outdata |>
  format_ae_specific(display = c("n", "prop", "diff"))
head(tbl$tbl)
```

Adding 95% confidence interval of M&N method. It requires to use `extend_ae_specific_inference`
to perform additional calculation.

```{r}
tbl <- outdata |>
  extend_ae_specific_inference() |>
  format_ae_specific(display = c("n", "prop", "diff", "diff_ci"))
head(tbl$tbl)
```

Adding average duration of AE. It requires to use `extend_ae_specific_duration`
to perform additional calculation.

```{r}
tbl <- outdata |>
  extend_ae_specific_duration(duration_var = "ADURN") |>
  format_ae_specific(display = c("n", "prop", "dur"))
head(tbl$tbl)
```

Adding average number of AE per subject. It requires to use `extend_ae_specific_events`
to perform additional calculation.

```{r}
tbl <- outdata |>
  extend_ae_specific_events() |>
  format_ae_specific(display = c("n", "prop", "events"))
head(tbl$tbl)
```

### Mock data preparation

We can also use `format_ae_specific` to create mock output data.

The purpose of the `mock` argument is not to create a comprehensive mock table template,
but a handy way to help user create a mock table that mimic the exact output layout.

Additional work is required to develop a flexible mock table generation tool
(e.g. a dedicated mock table generation package)

```{r}
tbl <- outdata |>
  format_ae_specific(mock = TRUE)
head(tbl$tbl)
```

### RTF tables

By using `tlf_ae_specific`, we can transfer the output from `format_ae_specific` to an RTF table.

```{r}
outdata |>
  format_ae_specific() |>
  tlf_ae_specific(
    meddra_version = "24.0",
    source = "Source:  [CDISCpilot: adam-adsl; adae]",
    path_outtable = "outtable/ae0specific1.rtf"
  )
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("outtable/ae0specific1.pdf")
```

The `tlf_ae_specific` function also provide some commonly used argument to customize the table.

```{r}
outdata |>
  format_ae_specific() |>
  tlf_ae_specific(
    meddra_version = "24.0",
    source = "Source:  [CDISCpilot: adam-adsl; adae]",
    col_rel_width = c(6, rep(1, 8)),
    text_font_size = 8,
    orientation = "landscape",
    path_outtable = "outtable/ae0specific2.rtf"
  )
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("outtable/ae0specific2.pdf")
```

We can also generate the mock table

```{r}
outdata |>
  format_ae_specific(mock = TRUE) |>
  tlf_ae_specific(
    meddra_version = "24.0",
    source = "Source:  [CDISCpilot: adam-adsl; adae]",
    path_outtable = "outtable/mock_ae0specific1.rtf"
  )
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("outtable/mock_ae0specific1.pdf")
```

## Future development

### Features

- Cross check the capability in SAS standard macro to identify required features.
