---
title: "Exposure Adjusted Event Rate"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Exposure Adjusted Event Rate}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This is a supplementary document for the file "Exposure-Adjusted Adverse Event
Summary (Updated)".
We first provide an explanation for the calculation of the exposure adjusted
event rate (EAER) of adverse events, and then briefly summarize the
programming steps of this calculation.

## EAER formula explanation

### EAER formula

$$
\begin{aligned}
EAER_j (\text{EAER for } Trt_j)
&= \frac{\text{total number of events for } Trt_j}{\text{total person-days for } Trt_j/(\text{exp factor})}  \\
&= \frac{\text{total number of events for } Trt_j \times \text{exp factor}}{\text{total person-days for } Trt_j}
\end{aligned}
$$

The exposure factor (exp factor) will be adjusted depending on the adjustment
unit defined by users. For instance, when the adjustment unit is
'100 person-month', EAER will be computed as follows:

$$
\begin{aligned}
EAER_j (\text{100 person-months})
&= \frac{\text{total number of events for } Trt_j \times \text{exp factor} (=100\times30.4367)}{\text{total person-days for } Trt_j}  \\
&= \frac{\text{total number of events for } Trt_j \times 3043.67}{\text{total person-days for } Trt_j}
\end{aligned}
$$

### EAER for different types of AEs

As an example, below are the EAER definitions for three types of AEs using
three treatment groups (PBO, Low Dose, High Dose).

#### ANY AE adj rate

$$
EAER_{PBO} (\text{100 person-months}) =\frac{\text{total number of AEs for PBO} \times 3043.67}{\text{total person-days for PBO}}
$$

$$
EAER_{LD} (\text{100 person-months}) =\frac{\text{total number of AEs for Low Dose} \times 3043.67}{\text{total person-days for Low Dose}}
$$

$$
EAER_{HD} (\text{100 person-months}) =\frac{\text{total number of AEs for High Dose} \times 3043.67}{\text{total person-days for High Dose}}
$$

#### Serious AE adj rate

$$
EAER_{PBO} (\text{100 person-months}) =\frac{\text{total number of SAEs for PBO} \times 3043.67}{\text{total person-days for PBO}}
$$

$$
EAER_{LD} (\text{100 person-months}) =\frac{\text{total number of SAEs for Low Dose} \times 3043.67}{\text{total person-days for Low Dose}}
$$

$$
EAER_{HD} (\text{100 person-months}) =\frac{\text{total number of SAEs for High Dose} \times 3043.67}{\text{total person-days for High Dose}}
$$

#### REL AE adj rate

$$
EAER_{PBO} (\text{100 person-months}) =\frac{\text{total number of Drug-Related AEs for PBO} \times 3043.67}{\text{total person-days for PBO}}
$$

$$
EAER_{LD} (\text{100 person-months}) =\frac{\text{total number of Drug-Related AEs for Low Dose} \times 3043.67}{\text{total person-days for Low Dose}}
$$

$$
EAER_{HD} (\text{100 person-months}) =\frac{\text{total number of Drug-Related AEs for High Dose} \times 3043.67}{\text{total person-days for High Dose}}
$$

## Programming steps of EAER

- Data preprocessing: At the end of this step, `adae` is created.
- Build a metadata object: At the end of this step, `meta` is created.
- Call the function `prepare_ae_exp_adj()`. This function has the following
  arguments: `meta`, `population`, `observation`, `parameter`, and `adj_unit`.
  The output will be a list:

```{r}
library(metalite.ae)
```

```{r}
meta <- meta_ae_example()

x <- prepare_ae_exp_adj(
  meta,
  population = "apat",
  observation = "wk12",
  parameter = "any;rel;ser",
  adj_unit = "month"
)

x
```

Run `x$adj_rate` to get the EAER:

```{r}
x$adj_rate
```
