---
title: "Introduction to metalite.ae"
output:
  rmarkdown::html_vignette
resource_files:
   - package_manual.pdf
   - outtable/*.pdf
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Introduction to metalite.ae}
---

```{r, echo = FALSE, message = FALSE, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = TRUE,
  comment = "#>",
  out.width = "100%",
  dpi = 150
)
```

```{r, include=FALSE}
library(metalite)
library(metalite.ae)
```

# Overview

`metalite.ae` is an R package for standard adverse events analysis including: 

- AE summary 
- Specific AE analysis 
- Specific AE subgroup analysis (TODO)
- AE listing (TODO)

## Workflow 

The general workflow is splited into three parts. 

1. define meta data information using `metalite` package 
1. prepare outdata using `prepare_xxx` functions
1. extend outdata using `extend_xxx` functions
1. format outdata using `format_xxx` functions
1. create TLFs using `tlf_xxx` functions 

For example, we can create a simple AE summary table as below. 

```{r}
meta_ae_dummy() |>
  prepare_ae_summary(population = "apat", 
                     observation = "wk12", 
                     parameter = "any;rel;ser") |>
  
  format_ae_summary() |>
  
  tlf_ae_summary(source = "Source:  [CDISCpilot: adam-adsl; adae]",
                 path_outtable = "outtable/ae0summary.rtf") 
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("outtable/ae0summary.pdf")
```

More examples can be found in [ae-summary](ae-summary.html) and [ae-specific](ae-specific.html) pages. 

## Highlighted features

- enable metadata structure 
- consistent input and output in standard functions
- streamline mock table generation








