---
title: "Introduction to metalite.ae"
output:
  rmarkdown::html_vignette
resource_files:
   - outtable/*.pdf
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Introduction to metalite.ae}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  out.width = "100%",
  dpi = 150,
  eval = TRUE
)
```

```{r, include=FALSE}
library(metalite)
library(metalite.ae)
```

## Overview

metalite.ae is an R package for standard adverse events analysis, including:

- AE summary.
- Specific AE analysis.
- Specific AE subgroup analysis (TODO).
- AE listing (TODO).

### Workflow

The general workflow is split into three parts.

1. Define metadata information using the metalite package.
1. Prepare outdata using `prepare_*()` functions.
1. Extend outdata using `extend_*()` functions.
1. Format outdata using `format_*()` functions.
1. Create TLFs using `tlf_*()` functions.

For example, we can create a simple AE summary table as below.

```{r}
meta_ae_example() |>
  prepare_ae_summary(
    population = "apat",
    observation = "wk12",
    parameter = "any;rel;ser"
  ) |>
  format_ae_summary() |>
  tlf_ae_summary(
    source = "Source:  [CDISCpilot: adam-adsl; adae]",
    path_outtable = "outtable/ae0summary.rtf"
  )
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("outtable/ae0summary.pdf")
```

More examples can be found in `vignette("ae-summary")` and `vignette("ae-specific")`.

### Highlighted features

- Enables metadata structure.
- Consistent input and output in standard functions.
- Streamlines mock table generation.
